Formation and evolution of intermediate-mass ratio inspirals

### 2020-01-16 article.tex

\newcommand{\Log}{{\rm Log}}
\newcommand{\Ms}{{\rm M}_\odot}
\newcommand{\Zs}{{\rm Z}_\odot}
\newcommand{\au}{{\rm AU}}
\newcommand{\gw}{{\rm GW}}
\newcommand{\gc}{{\rm GC}}
\newcommand{\kl}{{\rm KL}}
\newcommand{\ibh}{{\rm IMBH}}
\newcommand{\imbh}{{\rm IMBH}}
\newcommand{\inn}{{\rm in}}
\newcommand{\out}{{\rm out}}
\newcommand{\bh}{{\rm BH}}
\newcommand{\bhu}{{\rm BH1}}
\newcommand{\bhd}{{\rm BH2}}
\newcommand{\ARGdf}{\texttt{ARGdf}}
\newcommand{\ARCHAIN}{\texttt{ARCHAIN}}

@@

%
% Conclusions: What are the implications of our answer?
% ======================================================
%
We study the formation and evolution of intermediate mass ratio inspirals (IMRIs) triggered by the interactions between stellar black holes and an intermediate black hole (IMBH) inhabiting. We show the probability for an IMRI to form is an increasing function of the IMBH mass, representing a viable channel for the development of gravitational wave sources that can be possibly observed with the future generation of detectors.
\end{abstract}

@@

Therefore, the phase space that characterises the triple's main properties can be dissected into three groups:
\begin{itemize}
    \item {\bf Inner binary (IMBH+BH)}: depends on the IMBH and BH masses, $M_\ibh$ and $M_\bhu$, the binary semi-major axis $a_\inn$ and its eccentricity $e_\inn$.
    \item {\bf Outer binary (IMBH+BH+BH)}: depends on the third BH mass, $M_\bhd$, semi-major axis $a_\out$, eccentricity $e_\out$.
    \item {\bf Host cluster}: we model the host cluster as a \cite{Deh93} sphere, characterized by total mass, $M_c$, scale radius, $r_c$, and the inner slope of the density profile $\gamma$.
\end{itemize}

@@

The host cluster is assumed to be described by a \cite{Deh93} sphere, characterized by 
total mass $M_c$, scale radius $r_c$ and slope of the density profile $\gamma$.
The cluster length scale $r_c$ is selected randomly between 0.2 and 1.0 pc. For the cluster density slope, we assume a flat distribution with upper limit $\gamma \leq 1$.
To assign clusters' mass we either assume a scaling relation between the IMBH and the cluster mass, $M_\ibh - M_c$, or between the IMBH mass and the cluster velocity dispersion, $M_\ibh-\sigma_c$.

In the first case, we take advantage of the scaling relations based on numerical models of IMBH formation and observations of putative IMBHs in globular clusters \citep{zwart02,Lutzgendorf13,AS16}. In particular, we adopt the scaling provided by \cite{AS16}, connecting the host cluster mass $M_c$ with the total ''dark'' mass, inhabiting the cluster's centre, comprised of either an IMBH or a sizable population of stellar BHs
\begin{equation}
\Log \left(\frac{M_c}{\Ms}\right) = \alpha \Log \left(\frac{M_\ibh}{\Ms}\right) + \beta.
\label{mcmibh}
\end{equation}
with $\alpha = 0.999 \pm 0.001$ and $\beta = 2.23 \pm 0.009$.

Similarly to super-massive BHs in galactic nuclei, we can define the IMBH influence radius $R_{\rm inf}$, which delimits the region of space where IMBH dominates dynamics. This quantity can be connected to the cluster structural properties, namely
 \begin{equation}
 R_{\rm inf} = \frac{GM_\ibh}{\sigma_c^2} = \frac{M_\ibh}{M_c}R_{\rm hm},
 \end{equation}
 where we substituted the cluster velocity dispersion $\sigma_c$ with its value calculated at the cluster half-mass radius $R_{\rm hm}$, which in turn can be expressed in terms of the length scale and slope as
\begin{equation}
R_{\rm hm} = r_c(2^{1/(3-\gamma)}-1)^{-1}.
\end{equation} 
Combining the two equations above lead to 
\begin{equation}
R_{\rm inf} = \mu r_c(2^{1/(3-\gamma)}-1)^{-1},
\end{equation}
where $\mu = M_\ibh/M_c = 0.0058$ via Equation \ref{mcmibh}, thus implying that the the IMBH sphere of influence depends only on the cluster scale length and density slope.

In the second case, instead, we assume that GCs central regions represent a downsized version of galactic nuclei, thus the IMBH mass can be connected to the velocity dispersion inverting the so-called $M_\ibh - \sigma_c$ relation as 
\begin{equation}
     \Log \left(\frac{\sigma_c}{200{\rm km~s}^{-1}}\right) = \frac{1}{\alpha}\left(\left(\frac{\Log M_\ibh}{\Ms}\right) - \beta\right) ,
\end{equation}
with $\alpha = 4.24\pm0.41$ and $\beta=8.12\pm 0.08$ \citep{gultekin09}. Additionally, we associate to the cluster randomly a concentration parameter $c$ between $10.7$ and $131.4$, a range of values typical of dense star clusters characterized by an adimensional potential well $W_0 = 6-9$ \citep{king62}. The cluster mass is then calculated as $M_c = 2\sigma_c^2 c R_c / G$.

@@

\begin{table*}
\begin{center}
\caption{Main properties of our models}
\begin{tabular}{cccccccccccc}
\hline
model   & $f(\Log M_\ibh)$ & $M_\ibh$ & $\delta_\bh$ & $M_\bh$ & $a_\inn$ & $P(e_\inn)$ & $a_\out$ & $P(e_\out)$ &$P(\cos(i))$ &Cluster scaling & $N_{\rm sim}$\\ 
        &             & $\Ms$    &            & $\Ms$   & $\au$    &          & $\au$    &         &  &relation &\\
\hline 
0 & const & $10^2-5\times 10^5$ & $-2.2$ & $10-30$ & $10-315$ & $e^2$ & $20-630$ & $e^2$ & const&$M_\ibh -M_c$ & 2000\\
1 & const & $10^2-5\times 10^5$ & $-2.2$ & $10-30$ & $10-315$ & $e^2$ & $20-630$ & $e^2$ & const&$M_\ibh -\sigma_c$ & 2000\\
2 & const & $10^2-5\times 10^5$ & $-2.2$ & $10-30$ & $10-315$ & $e^2$ & $630-1580$ & $e^2$ &const &$M_\ibh -M_c$ & 2000\\
\hline
\end{tabular}
\end{center}
\label{tab:t2}
\end{table*}

@@

\subsection{IMRIs formation channels and subsequent evolution}

The evolution of the IMBH-BH-BH triple immersed in the host cluster gravitational field can lead to very different outcomes, as shown in Figure \ref{fig:f3}:
\begin{itemize}
    \item {\bf Disrupted triple} The three-body interaction undergoes a chaotic phases that ultimately lead to the ejection of a stellar BHs, whereas the inner binary either preserves its original components or undergoes a component swap.
    \item {\bf Hierarchical triple.} The triple arranges in a hierarchical configuration, with the outer BHs perturbing secularly the evolution of the inner binary. In this case, the inner binary can develop Kozai-Lidov oscillations \citep{kozai62,lidov62} that can trigger its eccentricity to grow to values close to unity. 
    \item {\bf Unstable triple.} The inner and outer binary are in an unstable configuration, whose evolution is mostly dominated by chaos.
\end{itemize}

@@

In the second case, the triple arranges in a hierarchical configuration, whose evolution is dominated by secular perturbations. This mechanism, known as Kozai-Lidov \citep{kozai62,lidov62}, can induce periodic oscillations in the mutual inclination and the inner binary eccentricity. A parameter widely used to discern hierarchical triples is defined as \citep{Lithwick11,naoz11}
\begin{equation}
\epsilon_\kl = \frac{M_\imbh - m_\bh }{M_\imbh + m_\bh} \frac{a_{\inn}}{a_{\out}}\frac{e_{\out}}{(1-e_{\out}^2)},
\end{equation}
being the triple hierarchical if $|\epsilon| > 0.01$.

@@

In the case of an unstable triple, the inner and outer binaries have similar orbital properties and dynamics is mostly dominated by chaotic interactions. According to the \cite{mardling01} criterion, a triple is unstable if
\begin{equation}
    \frac{a_\out}{a_\inn} < \frac{2.8}{1-e_\out}\left(1-\frac{0.3i}{\pi}\right)\left[\frac{(1+q_\out)(1+e_\out}{\sqrt{1-e_\out}}\right]^{2/5}, 
\end{equation}
being $q_\out = m_\out/(m_\imbh + m_\bh)$.

In our models, we calculate $\epsilon_\kl$ and $a_\out/a_\inn$ at the end of the simulations, in order to identify to which category a triple belongs. The merger time for the inner binary is calculated through Equation \ref{peters} for models categorized as ``Disrupted'' or ``Unstable'', whereas we use Equation \ref{kozai} for those identified as ``Hierarchical''.

@@

In SET0, an IMRI develop in $212$ cases out of 2000, i.e. $f_{\rm mer} = 10.6\%$ of formation probability. Concerning the formation channel, $8\%$ of IMRIs formed from Disrupted, $66.5\%$ from Hierarchical, and $25.5\%$ from Unstable . Similarly, in SET1 we find $268$ IMRIs ($13.4\%$), $4.9\%$ of them formed from Disrupted, $67.9\%$ from Hierarchical triples, and $27.2\%$ from Unstable. All IMRIs formation probability for different channels are summarized in Table \ref{tab:t3}.

\begin{table}
    \centering
    \caption{IMRIs formation probability}
    \begin{tabular}{cccccc}
        \hline
        \hline 
        model& $f_{\rm mer}$ & $f_{\rm mer,D}$ & $f_{\rm mer,H}$ & $f_{\rm mer,U}$\\
             & $\%$& $\%$& $\%$& $\%$\\
        \hline
        SET0 & $10.6$ & $8.0$ & $66.5$ & $25.5$ \\
        SET1 & $13.4$ & $4.9$ & $67.9$ & $27.2$ \\
        \hline
    \end{tabular}
    \label{tab:t3}
\end{table}

@@

\begin{figure}
\centering
\includegraphics[width=8cm]{semiaxis}\\
\includegraphics[width=8cm]{ecce}
\caption{Top panel: semi-major axis distribution for the inner binary. The black steps refer to initial values for all the simulations performed, while the red filled steps refer only to merger binaries and final values. Bottom panel: the same as in top panel, but for the inner binary eccentricity.}
\label{fig:f4}
\end{figure}

@@

To explore the dependence between IMRI candidates eccentricity at formation and IMBH mass, we take the eccentricity directly from the last snapshot for models targeted as ``Disrupted'' and ``Unstable'', while for ``Hierarchical'' we record the maximum eccentricity. We then calculate the mean eccentricity value $\langle e_\gw\rangle$  in different mass bins, making the same for the subset of IMRI candidates, as shown in Figure \ref{fig:f4}. In general, it seems that, regardless of the set, IMRI candidates with masses below $5\times10^3\Ms$ have, on average, $\langle e_\gw\rangle$ larger compared to the whole simulations sample, while this trend reverses for heavier IMRIs.

The average eccentricity increases in the $10^2-5\times10^3\Ms$ mass range, peaking at around $500-1000\Ms$, and than decreases rapidly at increasing the IMBH mass.This implies that the evolution of the inner binary leads the eccentricity to achieve larger values in correspondence of lighter IMBHs.

We find that IMBHs with masses typical of star clusters $10^3<M_\ibh/\Ms<10^4$ are expected to form, on average, extremely eccentric IMRIs, while the eccentricity at formation attains more moderate values at IMBH masses typical of low-mass dwarf galaxies and nuclear clusters.

While affecting the eccentricity, the IMBH mass does not have a significant effect on the companion BH mass. Indeed, IMRIs light component masses follows the assumed BH mass distribution.

\begin{figure}
\centering
\includegraphics[width=8cm]{average_ecc_sets}\\
\caption{The average eccentricity ($\langle e_{\rm mer}\rangle$), calculated at the end of simulaitons in different IMBH mass bins, as a function of the IMBH mass and for different sets. We show the IMRIs average eccentricity SET0 (filled red points), SET1 (open blue squares), and SET2 (filled black points). The total average value is also marked for SET0 (dashed red line), SET1 (dotted blue line), and SET2 (dot-dashed black line). }
\label{fig:f5}
\end{figure}

In order to uncover what is the role played by the IMBH in determining IMRIs formation, we calculate the 
percentage of IMRIs formed in different mass bins. This quantity, namely $f_{\rm mer}$, represents {\it de facto} IMRIs formation probability. We find that $f_{\rm mer}$ follows quite a precise trend, regardless of the set, thus in what follows we focus on SET0 only, for clarity's sake. Figure \ref{fig:f5} shows the $f_{\rm mer}-M_\ibh$ relation. We differentiate between IMRIs forming in disrupted triples, in a hierarchical or unstable configuration. 
Our results suggest that an IMBH-BH-BH has a $10\%$ of probability to arrange in a hierarchical configuration, regardless of the IMBH mass or the GW timescale. IMRIs formation probability, instead, increases weakly with IMBH mass, increasing from $\sim 2\%$ for $M_\ibh \simeq 10^2$ to up to $15\%$ for heavy IMBH, $M_\ibh>10^5\Ms$.

@@

In the IMBH mass range $M_\ibh>10^3\Ms$, the IMRIs formation probability is well described by a simple powerlaw
\begin{equation}
f_{\rm mer} = \alpha \left(\frac{M_\ibh}{10^2\Ms}\right)^\beta,
\label{fmerg}
\end{equation}  
with $\alpha = (7\pm2)\times 10^{-3}$ and $\beta = 0.45\pm0.03$.

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{mergers.eps}
    \caption{Percentage for IMRIs formation as a function of IMBH mass. Red open squares shows all the models that are hierarchical at the end of the simulation. Open dark blue circles identify IMRIs. From larger to smaller circles, blue filled circles represent IMRIs formed from unstable, hierarchical, and disrupted triples.}
    \label{fig:f5}
\end{figure}

@@

{\bf Previous studies based on full consistent N-body simulations of star clusters containing IMBHs with various masses and general properties  \citep{konstantinidis13,haster16,leigh14, macleod16} have shown that binary formation involving an IMBH takes place at a rate of $R_{\rm bin} \sim 10^{-7}$ yr. Assuming typical ages for globular-like star clusters of $t_{\rm age}\sim 10^{10}$ yr and assuming that IMBH buildup occurs in the clusters early lifetime ($\lesssim 1$ Gyr), we can use the merger probability fitting function depicted in Equation \ref{fmerg} to infer IMRIs merger rate per typical cluster
\begin{equation}
    \Gamma(M_\ibh) = R_{\rm bin}t_{\rm age}f_{\rm mer} \sim 0.7\left(\frac{M_{\ibh}}{10^2\Ms}\right)^{0.45},
\end{equation}
ranging from $\Gamma_{\rm IMRI} = 7-50$ yr$^{-1}$ for $M_\ibh = 10^2-10^4\Ms$, respectively.

To put such a quantity in the context of MW-like galaxies, we need to make assumptions on the IMBH putative mass function. As recently suggested by Arca Sedda et al (in prep.), out of the 151 Galactic GCs currently known, around 35 might be harbouring, at present time, an IMBH with mass in the range $10^3-10^4\Ms$. They find that the mass distribution of these putative IMBHs can be described by a Maxwellian in the form 
\begin{equation}
    f(x_\ibh) = \frac{a}{\sigma\sqrt(2\pi)}\exp{\left(-\frac{\left(x_\ibh - \mu\right)^2}{ (2\sigma^2)}\right)},
\end{equation}
with $x_\ibh\equiv \Log M_\ibh$, and $a=0.14\pm0.3$ , $\mu = 4.01\pm0.09$ and $\sigma = 0.4\pm0.1 $.

Combining equations above, we can define a IMRI merger rate for MW-like galaxies as
\begin{equation}
    \Gamma_{\rm MW} = f_\ibh N_\gc \int \Gamma(M_\ibh) f(x_\ibh) {\rm d}M_\ibh,
\end{equation}
where $f_\ibh N_\gc$ is the number of GCs containing an IMBH. This equation has analytical solution
which has analytical solution
\begin{align}
 \frac{\Gamma_{\rm MW} (x_\ibh)}{f_\ibh N_\gc} =&   \frac{a}{2} (A + B\mu) {\rm erf}\left(\frac{x_\ibh-\mu}{\sigma\sqrt{2\pi}} \right) + \nonumber\\
                                & - \frac{a}{\sqrt(2\pi)}B \sigma \exp{\left(\frac{-\left(x_\ibh-\mu\right)^2}{2\sigma^2}\right)},
\end{align}
 the coefficients $A = \alpha -\beta \Log(100)$ and $B \equiv \beta$ are conveniently manipulated from Equation \ref{fmerg}.

Assuming a typical IMBH population spanning the mass range $10^2-10^4\Ms$, a number of GCs $N_\gc = 151$, a fraction $f_\ibh \simeq 35/151$ of GCs hosting an IMBH we can infer the MW-IMRI rate 
\begin{equation}
\Gamma_{\rm MW,IMRI} \simeq 4\times 10^{-3} {\rm MW}^{-1} {\rm yr}^{-1}
\end{equation}

}

\subsection{Serendipitous formation of stellar BH binaries around an IMBH}

@@

\begin{figure}
\centering 
\includegraphics[width=8cm]{kozai_151}
\caption{Top panel: time evolution of the merger timescale for model number 151, in which two BHs pair together while the IMBH acts as a perturber. Central panel: mutual inclination as a function of time for the BH-BH pair in simulation number 151. Bottom panel: ecentricity as a function of time for the BH-BH pair in simulation number 151.}
\label{F9}
\end{figure}

>>>>

\newcommand{\derd}{{\rm d}}
\newcommand{\Log}{{\rm Log}}
\newcommand{\Ms}{{\rm M}_\odot}
\newcommand{\Zs}{{\rm Z}_\odot}
\newcommand{\au}{{\rm AU}}
\newcommand{\gw}{{\rm GW}}
\newcommand{\gc}{{\rm GC}}
\newcommand{\kl}{{\rm KL}}
\newcommand{\ibh}{{\rm IMBH}}
\newcommand{\imbh}{{\rm IMBH}}
\newcommand{\imri}{{\rm IMRI}}
\newcommand{\inn}{{\rm in}}
\newcommand{\out}{{\rm out}}
\newcommand{\bh}{{\rm BH}}
\newcommand{\bhu}{{\rm BH1}}
\newcommand{\bhd}{{\rm BH2}}
\newcommand{\ARGdf}{\texttt{ARGdf}}
\newcommand{\ARCHAIN}{\texttt{ARCHAIN}}

@@

%
% Conclusions: What are the implications of our answer?
% ======================================================
%
We study the formation and evolution of intermediate mass ratio inspirals (IMRIs) triggered by the interactions between two stellar black holes (BHs) and an intermediate black hole (IMBH) inhabiting a globular cluster. We show that the probability for IMRIs formation is an increasing function of the IMBH mass that is well described by a power-law for IMBH masses above $M_\ibh = 10^3\Ms$. We find that IMRIs have a formation probability $\sim 15\%$, regardless of the assumptions made. Our results suggest that in low-density clusters these interactions lead to the formation of stable triples, a fraction of which arrange in a hierarchical configurations and undergo Kozai-Lidov oscillations. In dense clusters, instead, the most frequent configuration is unstable, with the IMBH and the closest BH being strongly affected by the chaotic perturbations exerted by the third BH. 
We calculate IMRIs horizon redshift for several ground- and space-based gravitational waves (GWs) observatories. Combining this information with our results and observational limits on the cosmological distribution of galaxies and star clusters, we infer the IMRI merger rate for different detectors. Assuming a 4 yr observation timespan, we infer a merger rate of 1-4 events yr$^{-1}$ for LIGO in the IMRIs mass range $100-250\Ms$ up to redshift $z=0.1-0.6$, 1-30 yr$^{-1}$ for LISA ($M_{\rm imri} > 10^3\Ms$) up to redshift $z=1.5$. The next generation of detectors will allow us to observe IMRI out to the peak of GC formation and even beyond, with an expected rate of $200-2000$ yr$^{-1}$ for ET ($M_{\rm imri} < 10^3\Ms$) and $250-4400$ yr$^{-1}$ for DECIGO ($M_{\rm imri} < 60 - 5\times 10^4 \Ms$). We show that a considerable number of sources can be observed are multiband thus can be observed with both low- and high-frequency detectors, especially in the case of future detectors. Although highly eccentric at formation, we find that only $\sim 1\%$ of IMRI will preserve a high eccentricity when entering the mHz frequency band. 
\end{abstract}

@@

Therefore, the phase space that characterises the triple's main properties can be dissected into three groups:
\begin{itemize}
    \item {\bf Inner binary (IMBH+BH)}: depends on the IMBH and BH masses, $M_\ibh$ and $M_\bhu$, the binary semi-major axis $a_\inn$ and its eccentricity $e_\inn$.
    \item {\bf Outer binary (IMBH+BH+BH)}: depends on the third BH mass, $M_\bhd$, semi-major axis $a_\out$, eccentricity $e_\out$.
    \item {\bf Host cluster}: we model the host cluster as a \cite{Deh93} sphere, characterized by total mass, $M_\gc$, scale radius, $r_\gc$, and the inner slope of the density profile $\gamma$.
\end{itemize}

@@

The host cluster is assumed to be described by a \cite{Deh93} sphere, characterized by 
total mass $M_\gc$, scale radius $r_\gc$ and slope of the density profile $\gamma$.
The cluster length scale $r_\gc$ is selected randomly between 0.2 and 1.0 pc. For the cluster density slope, we assume a flat distribution with upper limit $\gamma \leq 1$.
To assign clusters' mass we either assume a scaling relation between the IMBH and the cluster mass, $M_\ibh - M_\gc$, or between the IMBH mass and the cluster velocity dispersion, $M_\ibh-\sigma_\gc$.

In the first case, we take advantage of the scaling relations based on numerical models of IMBH formation and observations of putative IMBHs in globular clusters \citep{zwart02,Lutzgendorf13,AS16}. In particular, we adopt the scaling provided by \cite{AS16}, connecting the host cluster mass $M_\gc$ with the total ''dark'' mass, inhabiting the cluster's centre, comprised of either an IMBH or a sizable population of stellar BHs
\begin{equation}
\Log \left(\frac{M_\gc}{\Ms}\right) = \alpha \Log \left(\frac{M_\ibh}{\Ms}\right) + \beta.
\label{mcmibh}
\end{equation}
with $\alpha = 0.999 \pm 0.001$ and $\beta = 2.23 \pm 0.009$.

Similarly to super-massive BHs in galactic nuclei, we can define the IMBH influence radius $R_{\rm inf}$, which delimits the region of space where IMBH dominates dynamics. This quantity can be connected to the cluster structural properties, namely
 \begin{equation}
 R_{\rm inf} = \frac{GM_\ibh}{\sigma_\gc^2} = \frac{M_\ibh}{M_\gc}R_{\rm hm},
 \end{equation}
 where we substituted the cluster velocity dispersion $\sigma_\gc$ with its value calculated at the cluster half-mass radius $R_{\rm hm}$, which in turn can be expressed in terms of the length scale and slope as
\begin{equation}
R_{\rm hm} = r_\gc(2^{1/(3-\gamma)}-1)^{-1}.
\end{equation} 
Combining the two equations above lead to 
\begin{equation}
R_{\rm inf} = \mu r_\gc(2^{1/(3-\gamma)}-1)^{-1},
\end{equation}
where $\mu = M_\ibh/M_\gc = 0.0058$ via Equation \ref{mcmibh}, thus implying that the the IMBH sphere of influence depends only on the cluster scale length and density slope.

In the second case, instead, we assume that GCs central regions represent a downsized version of galactic nuclei, thus the IMBH mass can be connected to the velocity dispersion inverting the so-called $M_\ibh - \sigma_\gc$ relation as 
\begin{equation}
     \Log \left(\frac{\sigma_\gc}{200{\rm km~s}^{-1}}\right) = \frac{1}{\alpha}\left(\left(\frac{\Log M_\ibh}{\Ms}\right) - \beta\right) ,
\end{equation}
with $\alpha = 4.24\pm0.41$ and $\beta=8.12\pm 0.08$ \citep{gultekin09}. Additionally, we associate to the cluster randomly a concentration parameter $c$ between $10.7$ and $131.4$, a range of values typical of dense star clusters characterized by an adimensional potential well $W_0 = 6-9$ \citep{king62}. The cluster mass is then calculated as $M_\gc = 2\sigma_\gc^2 c r_\gc / G$.

@@

\begin{table*}
\begin{center}
\caption{Main properties of our models}
\begin{tabular}{cccccccccc}
\hline
model   & $M_\ibh$ & $\delta_\bh$ & $M_\bh$ & $a_\inn$ & $a_\out$ & $P(e)$ &$P(\cos(i))$ &Cluster scaling & $N_{\rm sim}$\\ 
        & $\Ms$    &            & $\Ms$   & $\au$    &  $\au$    &         &  &relation &\\
\hline 
0 & $10^2-5\times 10^5$ & $-2.2$ & $10-30$ & $10-315$ & $20-630$ & $2e$ & const&$M_\ibh -M_\gc$ & 2000\\
1 & $10^2-5\times 10^5$ & $-2.2$ & $10-30$ & $10-315$ & $20-630$ & $2e$ & const&$M_\ibh -\sigma_\gc$ & 2000\\
2 & $10^2-5\times 10^5$ & $-2.2$ & $10-30$ & $10-315$ & $630-1580$ & $2e$ &const &$M_\ibh -M_\gc$ & 2000\\
\hline
\end{tabular}
\begin{tablenotes}
\item Col. 1: model ID. Col. 2: mass of the IMBH, we adopt a sampling flat in logarithmic values. Col. 3: slope of the BH mass function adopted. Col. 4: range of masses adopted for stellar BHs. Col. 5-6: range of initial values for the inner/outer semimajor axis, we adopt a sampling flat in  logarithmic values. Col. 7-8: adopted distribution for initial eccentricity/inclination. Col. 9: adopted scaling to connect the IMBH and the cluster masses. Col. 10: Number of simulations performed.
\end{tablenotes}
\end{center}
\label{tab:t2}
\end{table*}

@@

\subsection{IMRIs formation channels and subsequent evolution}
\label{sec:imris}
The evolution of the IMBH-BH-BH triple immersed in the host cluster gravitational field can lead to very different outcomes, as shown in Figure \ref{fig:f3}:
\begin{itemize}
    \item {\bf Disrupted triple.} The three-body interaction undergoes a chaotic phases that ultimately lead to the ejection of a stellar BHs, whereas the inner binary either preserves its original components or undergoes a component swap.
    \item {\bf Stable triple.} The triple arranges in a configuration that satisfies the stability criterion for three-body systems. 
	\begin{itemize}
		\item {\bf Hierarchical triple.}  The triple arranges in a hierarchical configuration, with the outer BHs perturbing secularly the evolution of the inner binary. In this case, the inner binary can develop Kozai-Lidov oscillations \citep{kozai62,lidov62} that can trigger its eccentricity to grow to values close to unity. 
	\end{itemize}
    \item {\bf Unstable triple.} The inner and outer binary are in an unstable configuration, whose evolution is mostly dominated by chaos.
\end{itemize}

@@

In the second case, the triple is dynamically stable and the effects of perturbations arising from one of the component emerges secularly.  According to the \cite{mardling01} criterion, a triple is stable if
\begin{equation}
    \frac{a_\out}{a_\inn} > \frac{2.8}{1-e_\out}\left(1-\frac{0.3i}{\pi}\right)\left[\frac{(1+q_\out)(1+e_\out}{\sqrt{1-e_\out}}\right]^{2/5}, 
\label{stabcri}
\end{equation}
being $q_\out = m_\out/(m_\imbh + m_\bh)$.

If the triple arranges in a hierarchical configuration, its evolution can be dominated by secular perturbations known as Kozai-Lidov \citep{kozai62,lidov62}, which can induce periodic oscillations in the mutual inclination and the inner binary eccentricity. A parameter widely used to discern hierarchical triples is defined as \citep{Lithwick11,naoz11}
\begin{equation}
\epsilon_\kl = \frac{M_\imbh - m_\bh }{M_\imbh + m_\bh} \frac{a_{\inn}}{a_{\out}}\frac{e_{\out}}{(1-e_{\out}^2)},
\end{equation}
being the triple hierarchical if $|\epsilon| > 0.01$.

@@

In the case of an unstable triple, the inner and outer binaries have similar orbital properties and dynamics is mostly dominated by chaotic interactions.

In our models, we calculate $\epsilon_\kl$ and $a_\out/a_\inn$ at the end of the simulations, in order to identify to which category a triple belongs. 
In the case of models categorized as ``Disrupted'', the merger time is calculated simply via 
Equation \ref{peters}. For ``Hierarchical'' models, instead, we adopt Equation \ref{kozai}. Finally, for both ``Stable'' and ``Unstable'' models, the merger time is inferred as the average of the initial and final merger time and the same quantity evaluated in correspondence of the maximum eccentricity. Note that for ``Stable'' triples the merger time does not change dramatically over the simulated time.

@@

In SET0, an IMRI develop in $259$ cases out of 2000, i.e. $f_{\rm mer} = 12.95\%$ of formation probability. Concerning the formation channel, $6.6\%$ of IMRIs formed from Disrupted, $61.4\%$ from Stable, and $32\%$ from Hierarchical. Similarly, in SET1 we find $312$ IMRIs ($15.6\%$), $4.2\%$ of them formed from Disrupted, $62.8\%$ from Stable triples, and $33\%$ from Hierarchical. In SET2, we still find a similar merger probability ($f_{\rm mer} = 15.36\%$), but in this case the number of Hierarchical configurations outnumbers the other class of merger candidates covering the $83.7\%$ of the whole merger population. This is due to the fact that a larger outer semiaxis makes easier to satisfy the stability criterion in Equation \ref{stabcri}.
All IMRIs formation probability for different channels are summarized in Table \ref{tab:t3}.

\begin{table}
    \centering
    \caption{IMRIs formation probability}
    \begin{tabular}{cccccc}
        \hline
        \hline 
        model& $f_{\rm mer}$ & $f_{\rm mer,D}$ & $f_{\rm mer,U}$ & $f_{\rm mer,H}$\\
             & $\%$& $\%$& $\%$& $\%$\\
        \hline
		SET0 & $12.95$ & $6.6$ & $61.4$ & $32$ \\
		SET1 & $15.60$ & $4.2$ & $62.8$ & $33$ \\
		SET2 & $15.36$ & $2.3$ & $14$   & $83.7$ \\
        \hline
    \end{tabular}
    \label{tab:t3}
\end{table}

@@

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{semiaxis}\\
\includegraphics[width=\columnwidth]{ecce}
\caption{Top panel: semi-major axis distribution for the inner binary. The black steps refer to initial values for all the simulations performed, while the red filled steps refer only to merger binaries and final values. Bottom panel: the same as in top panel, but for the inner binary eccentricity.}
\label{fig:f4}
\end{figure}

@@

To explore the dependence between IMRI candidates eccentricity at formation and IMBH mass, we take the eccentricity directly from the last snapshot for models targeted as ``Disrupted'' and ``Unstable'', while for ``Hierarchical'' we record the maximum eccentricity. We then calculate the mean eccentricity value $\langle e_\gw\rangle$  in different mass bins, making the same for the subset of IMRI candidates, as shown in Figure \ref{fig:f4}. In general, it seems that, regardless of the set, IMRI candidates with masses below $5\times10^3\Ms$ have, on average, $\langle e_\gw\rangle$ larger compared to the whole simulations sample, while this trend reverses for heavier IMRIs. 
The average eccentricity increases in the $10^2-5\times10^3\Ms$ mass range, peaking at around $500-1000\Ms$, and than decreases rapidly at increasing the IMBH mass.This implies that the evolution of the inner binary leads the eccentricity to achieve larger values in correspondence of lighter IMBHs. We find that IMBHs with masses typical of star clusters $10^3<M_\ibh/\Ms<10^4$ are expected to form, on average, extremely eccentric IMRIs, while the eccentricity at formation attains more moderate values at IMBH masses typical of low-mass dwarf galaxies and nuclear clusters.
While affecting the eccentricity, the IMBH mass does not have a significant effect on the companion BH mass. Indeed, IMRIs light component masses follows the assumed BH mass distribution.

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{average_ecc_sets}\\
\caption{The average eccentricity ($\langle e_{\rm mer}\rangle$), calculated at the end of simulaitons in different IMBH mass bins, as a function of the IMBH mass and for different sets. We show the IMRIs average eccentricity SET0 (filled red points), SET1 (open blue squares), and SET2 (filled black points). The total average value is also marked for SET0 (dashed red line), SET1 (dotted blue line), and SET2 (dot-dashed black line). }
\label{fig:f5}
\end{figure}

The merger times associated with IMRIs show a broad distribution regardless of the models set considered. Figure \ref{fig:f6} shows the cumulative distribution of $t_\gw$ calculated as discussed in Section \ref{sec:imris}.

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{TGW_distribution}
\caption{Cumulative distribution of IMRI merger times for all the models investigated.}
\label{fig:f6}
\end{figure}

In order to uncover what is the role played by the IMBH in determining IMRIs formation, we calculate the percentage of IMRIs formed in different mass bins. This quantity, namely $f_{\rm mer}$, represents {\it de facto} IMRIs formation probability. We find that $f_{\rm mer}$ follows quite a precise trend, regardless of the set, thus in what follows we focus on SET0 only, for clarity's sake.

\begin{figure}
	\centering
    \includegraphics[width=\columnwidth]{mergers}
	\caption{IMRIs formation probability as a function of the IMBH mass for SET0. Red open squares shows all the models that are in an either stable or hierarchical configuration at the end of the simulation. Open dark blue circles identify IMRIs. From larger to smaller circles, blue filled circles represent IMRIs formed from unstable, hierarchical, and disrupted triples.}
	\label{fig:f7}
\end{figure}

Figure \ref{fig:f7} shows the $f_{\rm mer}-M_\ibh$ relation in the case of SET0. We differentiate between IMRIs forming in disrupted triples, in a hierarchical or unstable configuration. 
Our results suggest that an IMBH-BH-BH has a $10\%$ of probability to arrange in a hierarchical configuration, regardless of the IMBH mass or the GW timescale. IMRIs formation probability, instead, increases weakly with IMBH mass, increasing from $\sim 2\%$ for $M_\ibh \simeq 10^2$ to up to $15\%$ for heavy IMBH, $M_\ibh>10^5\Ms$.

@@

Figure \ref{fig:f8} shows how $f_{\rm mer}$ varies for the three set explored. In the IMBH mass range $M_\ibh>10^3\Ms$, this relation is well described by a simple powerlaw
\begin{equation}
f_{\rm mer} = \alpha \left(\frac{M_\ibh}{10^2\Ms}\right)^\beta,
\label{fmerg}
\end{equation}  
with $\alpha = (1.5\pm0.2)\times 10^{-2}$ and $\beta = 0.39\pm0.02$.

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{merger_sets.eps}
    \caption{IMRIs formation probability as a function of the IMBH mass for the three simulated sets.}
    \label{fig:f8}
\end{figure}

@@

\subsection{The serendipitous formation of stellar BH binaries around an IMBH}

@@

\begin{figure}
\centering 
\includegraphics[width=\columnwidth]{kozai_151}
\caption{Top panel: time evolution of the merger timescale for model number 151, in which two BHs pair together while the IMBH acts as a perturber. Central panel: mutual inclination as a function of time for the BH-BH pair in simulation number 151. Bottom panel: ecentricity as a function of time for the BH-BH pair in simulation number 151.}
\label{F9}
\end{figure}
